Backscatter Modulation Design for Symbiotic Radio Networks

### 2021-11-23 main.tex
:pencil: Update backscatter model

				\subfloat[Block diagram]{

@@

					\label{fi:block_diagram}

@@

				\subfloat[One-port network model]{

@@

					\label{fi:one_port_network_model}

@@

				\caption{For a passive tag, the rectifier and demodulator rely on the incident electromagnetic wave for energy harvesting and downlink information decoding, while the load-switcher manipulate the reradiated signal for backscatter modulation.}

@@

			Consider a bistatic backscatter system that consists of an excitation source, a dedicated reader, and a passive tag. The excitation source generates a carrier wave signal, the dedicated reader decodes the tag message, and the tag simultaneously harvests energy, backscatters its own message, and demodulates the downlink information if necessary. As shown in Fig.~\subref*{fi:block_diagram}, a typical passive tag consists of a scattering antenna, an energy harvester, a integrated receiver\footnote{For example, \cite{Kim2021a} prototyped a compact-size pulse position demodulator based on an envelope detector, which brings great potential to coordination, synchronization, and reflection pattern control.}, a load-switching modulator, and on-chip components (e.g., micro-controller, memory, and sensors). A portion of the impinging signal is absorbed by the tag while the remaining is backscattered to the space, as illustrated in Fig.~\subref*{fi:one_port_network_model}. According to Green's decomposition \cite{Hansen1989}, the backscattered signal can be decomposed into the \emph{structural mode} component and the \emph{antenna mode} component. The former is fixed and depends on the antenna geometry and material properties\footnote{We assume the structural mode reflection can be modeled as part of the environment multipath and covered by channel estimation \cite{Boyer2014}.}, while the latter is adjustable and depends on the mismatch of the antenna and load impedance. Hence, the equivalent reflection coefficient at tag state $m$ is defined as\footnote{We assume the linear backscatter model where $\Gamma_m$ is irrelevant to the incident electromagnetic field at the tag \cite{Dobkin2012}.}

@@

			\begin{remark}
				The reflection coefficient plays an important role in various network designs. For example, $\Gamma_m = 0$ (perfect matching) achieves maximum power transfer that is optimal for Wireless Power Transfer (WPT), $\lvert \Gamma_m \rvert = 1$ (perfect mismatching) achieves fully signal reflection that is optimal for Intelligent Reflecting Surface (IRS), and $\Gamma_m \ne \Gamma_{m'}$ (adjustable matching) enables backscatter modulation.
			\end{remark}

@@

		\begin{subsection}{Backscatter Modulation}
			Tags perform backscatter modulation by switching the load impedance between different states. For an $M$-ary Quadrature Amplitude Modulation (QAM) at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$, the reflection coefficient $\Gamma_m$ maps to the desired signal constellation point $c_m$ as \cite{Thomas2012a}
			\begin{equation}
				\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
				\label{eq:backscatter_modulation}
			\end{equation}
			where $\alpha \in [0,1]$ is the reflection efficiency at a given direction. For simplicity, we consider an $M$-ary Phase Shift Keying (PSK) with zero-offset (TODO), where the $m$-th constellation point is
			\begin{equation}
				c_m = \exp \left(j \frac{2 \pi m}{M}\right).
				\label{eq:mpsk}
			\end{equation}
			\begin{remark}
				For passive tags, the reflection efficiency $\alpha$ controls the tradeoff between the backscatter strength and harvestable power. Interestingly, when $\alpha = 1$, the reflection coefficient set $\{\Gamma_m\}$ of the $M$-PSK backscatter coincides with that of an ideal discrete $M$-state uniform IRS. The optimal strategy for the IRS is to choose one reflection state with probability \num{1} to boost the equivalent channel, while the optimal strategy for the modulator is to utilize all constellation points with equal probability. It inspires one to optimize the tag input distribution to jointly benefit the backscatter modulation and passive beamforming.
			\end{remark}
		\end{subsection}
	\end{section}

	\begin{section}{The Multi-Tag Case}
		\begin{subsection}{System Model}
			\begin{figure}[!t]
				\centering
				\def\svgwidth{0.9\columnwidth}
				\import{assets/}{symbiotic_radio.eps_tex}
				\caption{A single-user multi-tag symbiotic radio system.}
				\label{fi:symbiotic_radio}
			\end{figure}
			As shown in Fig.~\ref{fi:symbiotic_radio}, we propose a single-user (UE) multi-tag (TG) symbiotic radio network where the RF signal generated by the $Q$-antenna Access Point (AP) is shared by two coexisting systems. In the primary downlink system, the AP transmits information to the single-antenna user. In the secondary backscatter system, the AP acts as the carrier emitter, the $K$ nearby single-antenna tags modulate their information over the reradiated RF signals, and the user serves as the multi-tag backscatter reader. Denote the AP-UE direct channel as $\boldsymbol{h}_{\mathrm{D}}^H \in \mathbb{C}^{1 \times Q}$, the AP-TG $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$, the TG $k$-UE backward channel as $h_{\mathrm{B},k}$, and the cascaded forward-backward channel of tag $k$ as $\boldsymbol{h}_{\mathrm{C},k}^H \triangleq h_{\mathrm{B},k} \boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$. For simplicity, we consider a quasi-static block fading model where the channel coefficients remain constant within each coherence interval and vary independently over different coherence intervals, and assume the coherence interval $T$ is much longer than the backscatter symbol period $T_c$ and primary symbol period $T_s$. We also assume the direct channel and all cascaded channels can be successfully estimated and fed back to the AP.\footnote{Due to the lack of RF chains at the passive tag, accurate and efficient Channel State Information at the Transmitter (CSIT) acquisition can be challenging. One possible approach is that the AP sends pre-defined pilots, the tags respond in well-designed manners, and the user performs least-square estimation with feedbacks \cite{Bharadia2015,Yang2015b,Guo2019e}.} Since the tags need to physically switch the loads for backscatter modulation, they communicate at a much longer symbol period (and thus lower rates) than the primary system. As such, we assume the transitions of all tags are perfectly synchronized, and the backscatter symbol period satisfies $T_c = N T_s$ where $N \gg 1$ is a positive integer.

			Without loss of generality, we focus on the transmissions and detections during the interval of one particular backscatter symbol. To provide a preliminary insight, we assume the primary symbol $s[n]$ at block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is in standard CSCG distribution, and the backscatter symbol $c_k$ of tag $k$ employs $M$-PSK modulation by \eqref{eq:mpsk}, $\forall k \in \mathcal{K}$. Thus, the signal received by the user at block $n$ can be expressed as\footnote{We omit the signal reflected by two or more times and assume the time difference of arrival from different paths are negligible\cite{Wu2019,Guo2019b}.}
			\begin{equation}
				y[n] = \left(\boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H c_k\right) \boldsymbol{w} s[n] + w[n],
				\label{eq:received_signal}
			\end{equation}
			where $\boldsymbol{w} \in \mathbb{C}^{Q \times 1}$ is the active precoder satisfying $\lVert \boldsymbol{w} \rVert^2 \le P$, $P$ is the average transmit power constraint at the AP, and $w[n] \sim \mathcal{CN}(0,\sigma^2)$ is the Additive White Gaussian Noise (AWGN) at block $n$. Besides, we stack the backscatter symbol of $K$ tags as $\boldsymbol{c} \triangleq [c_1,\ldots,c_K]^T \in \mathbb{C}^{K \times 1}$, the received signal over $N$ blocks as $\boldsymbol{y} \triangleq [y[1],\ldots,y[N]]^T \in \mathbb{C}^{N \times 1}$, and define the equivalent channel for primary transmission as
			\begin{equation}
				h_{\mathrm{E}}^H(\boldsymbol{c}) \triangleq \boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H c_k.
				\label{eq:equivalent_channel}
			\end{equation}

			\begin{remark}
				The proposed symbiotic radio system is reminiscent of a Multiple Access Channel (MAC) where the AP and the tags transmit simultaneously to the user. It inspired all existing papers [TODO] to perform Successive Interference Cancellation (SIC) that first non-coherently detects the primary message under backscatter uncertainty (requires non-coherent codes), then cancels its contribution and decodes the backscatter messages. However, different from the conventional MAC with Superposition Coding (SC), the symbiotic radio system involves Multiplication Coding (MC) because the primary and secondary messages are indeed combined in the received signal by multiplication. Also, the computational complexity of SIC can be high since it requires $K$ re-encoding, precoding, and subtraction operations. Therefore, novel multi-stream detection techniques should be tailored for symbiotic radio systems.
			\end{remark}
		\end{subsection}

		\begin{subsection}{Signal Detection}
			The main difference between symbiotic radio and conventional MAC is that the primary message also reaches the user from the backscatter link. This characteristic inspires one to first decode the tag message, then model its contribution within the equivalent channel during primary decoding (i.e., unify backscatter decoding and channel training), instead of performing SIC. In such case, the reflection pattern not only embeds the tag message but also adjusts the legacy channel in a controllable manner. Therefore, for a fixed fading block, the primary transmission is able to achieve the ergodic capacity with artificial channel variation created by the backscatter modulation. [Channel over channel, randomness over randomness]

>>>>

				\subfloat[Block diagram of a passive tag]{

@@

					\label{fi:block_diagram_of_a_passive_tag}

@@

				\subfloat[Backscatter modualtion]{

@@

					\label{fi:backscatter_modualtion}

@@

				\caption{For a passive tag, the rectifier and demodulator rely on the incident electromagnetic wave for energy harvesting and source decoding, while the load-switcher adjusts the reradiated signal for backscatter modulation.}

@@

			A bistatic backscatter system consists of an excitation source, multiple (semi-)passive tags, and a backscatter reader. When illuminated, the tags simultaneously harvest energy, backscatter message, and demodulate the source signal if necessary. Fig.~\subref*{fi:block_diagram_of_a_passive_tag} shows a typical passive with a scattering antenna, an energy harvester, an integrated receiver\footnote{For example, \cite{Kim2021a} prototyped a compact-size pulse position demodulator based on an envelope detector, which brings great potential to coordination, synchronization, and reflection pattern control.}, a load-switching modulator, and on-chip components (e.g., micro-controller, memory, and sensors). A portion of the impinging signal is absorbed by the tag while the remaining is backscattered to the space. According to Green's decomposition \cite{Hansen1989}, the backscattered signal can be decomposed into the \emph{structural mode} component and the \emph{antenna mode} component. The former is fixed and depends on the antenna geometry and material properties\footnote{The structural mode component can be regarded as part of the environment multipath and modeled by channel estimation \cite{Boyer2014}.}, while the latter is adjustable and depends on the mismatch of the antenna and load impedance. Fig.~\subref*{fi:backscatter_modualtion} illustrates a simplified circuit and backscatter model at tag state $m$. The corresponding reflection coefficient is defined as\footnote{We assume the linear backscatter model where the reflection coefficient is irrelevant to the incident electromagnetic field at the tag \cite{Dobkin2012}.}

@@

@@

		\begin{subsection}{Backscatter Modulation}
			Backscatter modulation is achieved by switching the tag load impedance between different states. For an $M$-ary Quadrature Amplitude Modulation (QAM) at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$, the reflection coefficient $\Gamma_m$ maps to the signal constellation point $c_m$ as \cite{Thomas2012a}
			\begin{equation}
				\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
				\label{eq:backscatter_modulation}
			\end{equation}
			where $\alpha \in [0,1]$ is the reflection efficiency at a given direction. For simplicity, we consider an $M$-ary Phase Shift Keying (PSK) with zero-offset (TODO), where the $m$-th constellation point is
			\begin{equation}
				c_m = \exp \left(j \frac{2 \pi m}{M}\right).
				\label{eq:mpsk}
			\end{equation}
			\begin{remark}
				For passive tags, the reflection efficiency $\alpha$ controls the tradeoff between the harvestable power and backscatter strength: $\alpha = 0$ corresponds to maximum power transfer to the tag, while $\alpha = 1$ with $M$-PSK corresponds to ideal Intelligent Reflecting Surface (IRS) with $M$ discrete states.
			\end{remark}
		\end{subsection}
	\end{section}

	\begin{section}{The Multi-Tag Case}
		\begin{subsection}{System Model}
			\begin{figure}[!t]
				\centering
				\def\svgwidth{0.9\columnwidth}
				\import{assets/}{symbiotic_radio.eps_tex}
				\caption{A single-user multi-tag symbiotic radio system.}
				\label{fi:symbiotic_radio}
			\end{figure}
			As shown in Fig.~\ref{fi:symbiotic_radio}, we propose a single-user (UE) multi-tag (TG) symbiotic radio network where the RF signal generated by the $Q$-antenna Access Point (AP) is shared by two coexisting systems. In the primary downlink system, the AP transmits information to the single-antenna user. In the secondary backscatter system, the AP acts as the carrier emitter, the $K$ nearby single-antenna tags modulate their information over the reradiated RF signals, and the user serves as the multi-tag backscatter reader. Denote the AP-UE direct channel as $\boldsymbol{h}_{\mathrm{D}}^H \in \mathbb{C}^{1 \times Q}$, the AP-TG $k \in \mathcal{K} \triangleq \{1,\ldots,K\}$ forward channel as $\boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$, the TG $k$-UE backward channel as $h_{\mathrm{B},k}$, and the cascaded forward-backward channel of tag $k$ as $\boldsymbol{h}_{\mathrm{C},k}^H \triangleq h_{\mathrm{B},k} \boldsymbol{h}_{\mathrm{F},k}^H \in \mathbb{C}^{1 \times Q}$. For simplicity, we consider a quasi-static block fading model where the channel coefficients remain constant within each coherence interval and vary independently over different coherence intervals, and assume the coherence interval $T$ is much longer than the backscatter symbol period $T_c$ and primary symbol period $T_s$. We also assume the direct channel and all cascaded channels can be successfully estimated and fed back to the AP.\footnote{Due to the lack of RF chains at the passive tag, accurate and efficient Channel State Information at the Transmitter (CSIT) acquisition can be challenging. One possible approach is that the AP sends pre-defined pilots, the tags respond in well-designed manners, and the user performs least-square estimation with feedbacks \cite{Bharadia2015,Yang2015b,Guo2019e}.} Since the tags need to physically switch the loads for backscatter modulation, they communicate at a much longer symbol period (and lower rates) than the primary system. As such, we assume the transitions of all tags are perfectly synchronized, and the backscatter symbol period satisfies $T_c = N T_s$ where $N \gg 1$ is a positive integer.

			Without loss of generality, we focus on the transmissions and detections during one particular backscatter symbol period. To provide a preliminary insight, we assume the primary symbol $s[n]$ at block $n \in \mathcal{N} \triangleq \{1,\ldots,N\}$ is in standard CSCG distribution, and the backscatter symbol $c_k$ of tag $k$ employs $M$-PSK modulation by \eqref{eq:mpsk}, $\forall k \in \mathcal{K}$. Thus, the signal received by the user at block $n$ can be expressed as\footnote{We omit the signal reflected by two or more times\cite{Wu2019} and assume the time difference of arrival from different paths are negligible\cite{Guo2019b}.}
			\begin{equation}
				y[n] = \left(\boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H c_k\right) \boldsymbol{w} s[n] + w[n],
				\label{eq:received_signal}
			\end{equation}
			where $\boldsymbol{w} \in \mathbb{C}^{Q \times 1}$ is the active precoder satisfying $\lVert \boldsymbol{w} \rVert^2 \le P$, $P$ is the average transmit power constraint at the AP, and $w[n] \sim \mathcal{CN}(0,\sigma^2)$ is the equivalent Additive White Gaussian Noise (AWGN) at block $n$. Besides, we stack the backscatter symbol of $K$ tags as $\boldsymbol{c} \triangleq [c_1,\ldots,c_K]^T \in \mathbb{C}^{K \times 1}$, the received signal over $N$ blocks as $\boldsymbol{y} \triangleq [y[1],\ldots,y[N]]^T \in \mathbb{C}^{N \times 1}$, and define the equivalent channel for primary transmission as
			\begin{equation}
				h_{\mathrm{E}}^H(\boldsymbol{c}) \triangleq \boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H c_k.
				\label{eq:equivalent_channel}
			\end{equation}

			\begin{remark}
				The proposed symbiotic radio system includes a multiplicative Multiple Access Channel (MAC) where the AP and the tags simultaneously transmit to the user. It inspired [TODO] to perform Successive Interference Cancellation (SIC) that first non-coherently detects the primary message under backscatter uncertainty, then cancels its contribution and decodes the tag messages. This scheme requires non-coherent coding at the AP and $K$ re-encoding, precoding, and subtraction operations at the user. However, different from the conventional MAC with Superposition Coding (SC), the symbiotic radio system involves Multiplication Coding (MC) that combines the primary and secondary messages by multiplication. Hence, novel multi-stream detection techniques should be tailored to symbiotic radio systems to accommodate the massive connectivity of tags and the multiplicative combination of links.
			\end{remark}
		\end{subsection}

		\begin{subsection}{Signal Detection}
			Based on a legacy downlink receiver, we propose a low-complexity multi-tag energy detection scheme in the presence of primary interference. It is achieved by extending the AmBC detection to the case of multiple sources.
			\begin{remark}
				One key property of symbiotic radio is the primary message propagates to the user from a known channel and multiple multiplicative channels with uncertainty introduced by backscatter modulation. If the backscatter symbols can be successfully decoded first, they can be modeled within the equivalent channel \eqref{eq:equivalent_channel} as channel training, instead of being removed by SIC. As such, the reflection coefficient simultaneously encodes the tag message and influences the equivalent channel for legacy transmission.
			\end{remark}
